\documentclass[upright, contnum]{umemoria}

\usepackage[lined,ruled]{algorithm2e}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{ragged2e}
\usetikzlibrary{shapes}

% allow full width tables
\usepackage{tabularx}

\usepackage{multirow}

% to do notes
\usepackage{todonotes}
 
% spacing
\usepackage{setspace}

% colors
\usepackage{xcolor}

% bold math
\usepackage{bm}

% line numbers 
\usepackage[switch]{lineno}
\renewcommand{\linenumberfont}{\normalfont\scriptsize\color{red}}
%\linenumbers

% boxes
\usepackage{tcolorbox}

% algorithms
\usepackage{algorithmic}

% tablas de high-activity
\usepackage{pbox}

% tablas de mas de una pagina
\usepackage{longtable}

%%%% custom vars

% vector characters
\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}

% T for transpose
\newcommand{\customT}{\text{T}}

% TP, FN, etc..
\newcommand{\cTP}{\text{TP}}
\newcommand{\cTN}{\text{TN}}
\newcommand{\cFP}{\text{FP}}
\newcommand{\cFN}{\text{FN}}

%fix for the oneside argument
\makeatletter
\g@addto@macro\titlepage{\pagenumbering{Alph}}
\g@addto@macro\endtitlepage{\pagenumbering{roman}}
\makeatother

\depto{Departamento de Ciencias de la Computación}
\author{Mauricio Daniel Quezada Veas}
\title{\bf KNOWLEDGE DISCOVERY FROM NEWS EVENTS ON TWITTER}
\auspicio{\break CONICYT PCHA/Doctorado Nacional 2015/21151445 y \break por el Instituto Milenio de Fundamentos de los Datos, IMFD}
\date{Marzo 2020}
\guia{BÁRBARA POBLETE LABRA}
\carrera{Doctor en Ciencias, Mención Computación}
\memoria{Tesis para optar al Grado de \break  Doctor en Ciencias, Mención Computación}
\comision{AIDAN HOGAN \break MARCELO MENDOZA \break BRIAN D. DAVISON}

\usepackage{lipsum}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
 
\begin{document}

\frontmatter
\maketitle 

\begin{abstract} 

    %\todo[inline]{revisar el abstract}
    Online activity on social media involves the consumption and production of
    news event-related content. 
    %
    There are about 500 million Twitter messages published every day, and
    according to surveys, 59\% of its users use the platform as a way to obtain
    news. 
    %
    The high rate at which users produce multimodal content (text, images, and
    videos) requires flexible models to understand the dynamics of the
    information disseminated on social media. 
    %
    This thesis proposes the creation of context-aware models of events based on
    user-generated content in online social networks.
    %
    The goal of these models is to provide a means to perform high-level
    quantitative analysis of news events. 
    %
    The proposed models are designed from three perspectives: 
    %
    (1) the activity of users that react to an event when it happens, 
    %
    (2) the spatio-temporal context in which events develop, and 
    %
    (3) the content aggregations that can be exploited to generate a
    comprehensive representation of an event. 
    %
    For this work, we collected a dataset of news-related tweets dating from
    2013 to 2015, encompassing twenty thousand events, corresponding to about
    two hundred million tweets.
    %

    %
    First, we observe that events have particular characteristics based on user
    activity, and these characteristics are quite distinguishable at the early
    stages of their outbreak.
    %
    We present an event representation model based on user activity,
    characterize events based on different levels of activity, and show that it
    is possible to predict with high precision the level of activity an event
    would have by using aggregated event features.
    % 
    Second, we then develop a spatio-temporal representation of events based on
    geopolitical entities (e.g., countries).
    %
    We characterize events based on the countries related to them, and show that
    this representation allows us to portray real-world relationships between
    countries based on social media content.
    %
    Finally, we study content features of events and introduce a lightweight
    representation of events based on common URLs shared in social media
    messages.
    %
    We suggest that aggregating content based on shared URLs allows us to
    identify event topics efficiently, without losing considerable content
    quality.
    %
    These different event representations are useful, novel, and effective to
    perform high-level quantitative analysis of news events.

\end{abstract}

%\begin{dedicatoria} % opcional
%Una dedicatoria corta. Por ejemplo, al \emph{Centro Tecnológico Ucampus}
%\end{dedicatoria}

%\begin{thanks} % opcional
%\lipsum[1-2]
%\end{thanks}
%\cleardoublepage

\tableofcontents
\listoftables % opcional
\listoffigures % opcional

\mainmatter

\include{chapters/intro/intro-main}
\include{chapters/background/background-main}
\include{chapters/related-work/related-work-main}
\include{chapters/data/data-main}
\include{chapters/high-impact/high-impact-main}
\include{chapters/geopolitical/geopolitical-main}
\include{chapters/url-model/url-model-main}
\include{chapters/conclusions/conclusions-main}

% \input{glosario.tex} % opcional

\bibliographystyle{plain}
\bibliography{bibliografia}

\include{chapters/appendices/appendices-main} % opcionales

%\listoftodos[Notes]  % TODO QUITAR ESTA LISTA !

\end{document}
