\begin{intro}

% WHAT IS SOCIAL MEDIA

The so-called {\em Web 2.0} represents a change of state in how users interact
with the Web.
%
It is mainly defined as a platform destined to encourage end-users to
publish content. 
%
One of the main manifestations of this phenomena are the {\em online social
networks}, also called {\em microblogging platforms}.
%
Users online make connections with others based on different criteria, and start
to produce content that might be interesting to other users. 
%
Microblogging services such as Facebook~\cite{facebook}, Twitter~\cite{twitter},
or Sina Weibo~\cite{weibo} are nowadays among the most used platforms for users
to connect with family, friends, acquaintances, co-workers, or strangers with
similar interests.   
%
Users interact with each other and produce or share content, which could be
about their lives, their thoughts, about what is happening around the world,
etc.    
%
This collective of information published in these Internet-based applications,
such as microblogging platforms, blogs, wikis, etc., is what is called {\em
social media}~\cite{kaplan2010users}.

%%

% CONTENT IN SM

Content in social media is multimodal. 
%
Twitter, for example, encourages users to publish short texts (initially limited
to 140 characters, now 280), but it recently started to incite users to share
more photographs~\cite{brown_2019}.
%  
And with the proliferation of smartphones and internet-connected devices, more
of this information is also geo-tagged and ``real-time''. 
%
Foursquare~\cite{foursquare} (and then
Swarm~\cite{swarm}) is an application for users to tag
and comment on locations around the world, which serves as a repository of data
about business and points of interest.
%
Whether via text, images, videos, sounds or hyperlinks, social media lowered the
entry barriers to content producers and made it easy for consumers to access to
a myriad of different pieces of information.


%%


% UTILITY, APPLICATIONS OF SM

The influence of social media on society can not be denied. 
% 
It has facilitated the communication and speeded-up the difussion of information
online. 
%
For instance, it is believed that the revolutionary wave of protests and
uprisings in the Arab states (known as the Arab Spring which began in 2010) was
highly influenced by social media as a means to organize and facilitate
communication~\cite{howard2011opening}. 
%
Also, for instance, it has permitted many applications in emergency management
and detection, such as earthquake alert systems using
Twitter~\cite{Sakaki2010,Sarmiento:2018:DDE:3201064.3201077,Mendoza2019}.
%
The usefulness of quantiative analysis of events through time is undeniable, and
social media offers a window to see and capture information about those events,
how they develop, and how the world interprets them.



% NEWS CONSUMPTION AND PRODUCTION


One of the main usages of social media platforms is the consumption and
generation of event-related content. 
%
According to a recent 2018 study~\cite{pewresearch}, about two thirds of U.S.
adults get their news on social media.
%
The most used platforms to get the news are Facebook, Youtube, and Twitter,
while over 70\% of Twitter users surveyed use that platform to do so.
%
Furthermore, nowadays almost every news outlet has a presence in social media,
in order to attract readers and viewers.
%
In this way, users comment on the news events, reacting to them according to a
myriad of factors, and many of these characteristics are present in one way or
another in social media.
%
Therefore, we see social media as a medium that reflects an important 
part of what society thinks about what is happening the world.




% PROBLEMS

However, the popularity of social media is not without issues. 
%
{\em Information overload} refers to the problem of being unable to manage or to
make decisions based on data, due to the high volume of information available
and the limited capabilities of the person who is dealing with it. 
%
Humans have limited cognitive processing capacities, and when they are
overloaded with information, their quality of decision making
suffers~\cite{gross1964managing}. 
%
In the context of social media, the high availability of diverse information may
prevent users to find relevant content.

%%

Finding relevant content in social media is not easily solved by search engines.
%
Publications on social media, or {\em posts}, can be of {\em varied quality}.
%
Posts are composed of multimedia pieces, but often they are brief.
%
For instance, a post can be a very short text, an hyperlink, a single image or
video with little context.
%
They can be also irrelevant to the user's interest, for example, spam posts,
which contain relevant keywords but in a misleading way, in order to lure users
into a irrelevant website. 
%
Posts can be also out-dated, delivering incorrect or obsolete information. 
%
Many posts can be duplicate ones, published by automated agents, or by users
using ``share buttons'' in websites which publish posts with a template text; 
%
they can be also near-duplicate, with little text differences, or sharing the
same resource from different URLs.
%
Another important characteristic of posts is that they are written in natural
language, so they can be incorrectly capitalized, mispelled, or with ambiguous
meaning.
%
Users make use of colloquial language and different forms of expression when
publishing content, e.g., abbreviations, hashtags (tags to describe content),
emojis (ideograms), etc.
%
Finally, messages can be also misleading, sharing false information.
%
All of these particularities of social media make it difficult to apply standard
techniques in order for users to find relevant content.

%%

At the same time, there is a massive scale of production of content. 
%
Twitter reports that there are about 320 million active monthly users and 500
million daily tweets published in their platform~\cite{twitter2014}. 
%
(These numbers were first stated in 2014 and have not been updated since then.)
%
Facebook has nearly 2.32 billion active monthly users~\cite{fbnewsroom}. 
%
Sina Weibo has 462 million active monthly users, while 93\% of them are on
mobile devices~\cite{chinawatch}.
%
The amount of content being published requires novel ways to deal with social
media data, in particular, regarding news events.


%%

We regard news events as a higher level abstraction than single
posts.
%
In related work, an event\footnote{In this work, we will use the term {\em news
event} and {\em event} indistinctly.} is deemed as {\em something that happens
in a certain place and time}~\cite{yang1999learning}, while other definitions
consider an event as a collection of documents related to a certain
occurrence~\cite{Becker:2010:LSM:1718487.1718524}.
%
Throughout this dissertation, we will consider an event as a collection of
social media posts describing or commenting on a real-world occurrence.
%
In this sense, an event is a more complex piece of information compared to
single posts as it leads to new tasks, such as event detection, tracking, or
summarization.
%
Also, an event is comprised of posts of heterogeneous quality, from different
locations, and at different times.
%
This yields to new problems and challenges when studying social media.

%%

% PROPOSAL

In this dissertation, we tackle the problem of {\em extracting useful knowledge}
from events on social media. 
%
In order to be able to infer and extract useful information from events, we
propose the development of event representations that leverage specific features
according to the desired goal when analyzing social media data.
%
For this, we propose different models or representations of events based on
three perspectives:
%

\begin{enumerate}
    \item {\bf User activity.} 
    %
    When users react to an event, they may manifest this reaction on social
    media, producing or sharing content relevant to the event. 
    %
    The characteristics of these manifestations are dependant on the proper
    features of the occurrence, and not all are equal. 
    %
    We look at how the activity of users can give us insights on the proper
    features of an event, and incorporate this behavior in a compact
    representation.

    \item {\bf Spatio-temporal context.} 
    %
    Events develop in different locations. 
    %
    On the other hand, users from different locations may react differently to
    the same event.
    % 
    We study the development of events based on user activity conditioned by the
    location users are from, proposing a representation of events and locations
    based on social media posts.

    \item {\bf Common features in content.} 
    %
    Users may publish similar pieces of content in social media in reaction to
    events.
    %
    However, each post can contribute to a different aspect of the event, while
    having some features in common.
    %
    We leverage these commonalities in content to produce a compact model that
    preserve topical information in events.
    
\end{enumerate}


% OBJETIVES

Our main objective is to develop event representations through different data
aggregations in order to perform quantiative analysis of news events on social
media.
%
Currently, it is very difficult to manage and analyze the high volume of
information being published when a event happens in the world. 
%
In particular, we want to study events through three perspectives: user reaction
and activity, spatio-temporal context of events, and content aggregation.
%
Understanding reaction involves discriminate which events are more important or
produce more impact in a community. 
%
Understanding spatio-temporal context refers to understand how communities from
different locations are affected by different events, as seen on social media,
and identify similar communities and events based on this context. 
%
Understanding content refers to the identification of the core aspects of an
event, without having to go through all the –potentially several– posts.
%
In particular, our goal is to propose different models for representing events. 
%
These models should be flexible enough to apply diverse methodologies to
discover useful knowledge from information published on social media about
real-world events, from the perspectives described above. 


%%

% WHY TWITTER

We chose Twitter as our data source for this work. 
%
The advantages of using Twitter are that, first, it provides a simple way to
obtain data and via its API (Application Programming Interface).
%
Second, it is not as restrictive as other sources, such as Facebook, which
incentivate users to mantain a private profile, hence making the data collection
much more difficult or impossible to perform.
%
Also, services like Facebook encourage users to share potentially personal
content, and not only event-related.
%
Third, Twitter it is primarily dedicated to incentivate sharing event-related
posts, and it is mainly used as a news source by its users (about 70\% of its
users use Twitter to get news). 
%
For instance, its website asks “What’s happening?” to users when publishing
content, as opposed to Facebook’s “What’s in your mind?”
%
\dots




\paragraph{Thesis statement.} 
%
This dissertation poses to define flexible models for events on the social
networking platform Twitter. 
%
Having three perspectives in mind, user reaction, spatio-temporal context, and
content, the defined models should be able to allow us to discover new insights
about news events reflected on Twitter. 
%
In particular, each perspective approaches specific Data Mining and Information
Retrieval techniques: 
%
the study of context involves modeling and exploratory data analysis; 
%
reaction involves filtering and classification;
%
content involves topic detection and document modeling.


The thesis statement is as follows:

\begin{center}
{\em
Modeling news events from user-contributed content on Twitter, based on
their spatio-temporal context, the reaction the users had on them, or the
multimedia content which the events contain, is novel and effective to perform
high-level quantitative analysis of news events.
}
\end{center}



\paragraph{Challenges.}




\paragraph{Contributions.}

\begin{enumerate}
\item an event representation based on user activity triggered by the news;
\item a spatio-temporal representation based on the location where an event
happens, and the locations the users commenting on the news are from; and
\item a compact representation of content based on shared URLs.
\end{enumerate}

In the context of events, it is often
difficult to comprehend what is happening in the world due to the high volume of
available information, potentially redundant, or of little relevance. On the
other hand, it is very difficult to assess the different aspects of an event,
when one of these aspects usually take high importance and the others are
relegated to be less important or popular, and hence more difficult to find
related information.
%%


The study of news events on social media has several applications in the
proposed setting. How the community reacts to different events would allow us to
identify the characteristics of these events by a measure of reaction or other
features given by the social network or the content of the events. By these
characteristics, it would be possible to identify or even predict which events
are going to cause a significative reaction from the community, improving
journalistic coverage or better response from authorities facing an emergency.
Additionally, by studying not only the response, but the context of different
communities and how they respond to certain events, may give us insights about
the communities themselves, for example, by revealing unexpected relations
between different communities, or by measuring event similarity using the
context, instead of content-based features. On the other hand, the study of the
content is useful to understand the different points of view ahead of an event.
Users accustomed to the same perspectives given by other users or sources may be
oblivious of other angles of the same news event. Being capable of identify the
different aspects of an event and then present these aspects in a concise
summary can deal with this problem. All in all, the proposed framework can be of
utility to understand social behavior, to study and decrease the effects of the
information overload and the echo chamber effect, as well as to perform
comparative historical research.





The echo chamber effect (also known as the filter bubble) refers to the
situation when, in the end, one’s own opinions are reflected back to oneself.
Posts and articles are customized to each individual, via recommendation [57],
due to media coverage [48], or by “the tendency of individuals to associate and
bond with similar others” 1 , (homophily, which has been seen and studied on
online social networks [39, 19]), and hence being exposed to similar opinions.
This effect produces polarization and potentially extreme bias to certain
tendencies or opinions. 


Comparative historical research 2 is a method to examine
historical events to create explanations about long-term tendencies or social
processes. So far, historical research has been restricted to traditional
archival data and historians’ written account of past events. Nevertheless, it
is undeniable that the data poured into social media about events is of great
value to society.







Here, a news event is defined as an abstraction of social media data. An event
is represented as a conglomerate of all the retrieved user-generated posts about
a real-world occurrence. It has been observed that using aggregated features
from several posts is more effective to reduce the variance in several studies
about social media. This is followed by the wisdom of the crowds principle,
which argues that the aggregation of information from multiple users results in
better decisions than done by single members of the group. Next we specify what
we mean by context, content and reaction in the thesis proposal:

Spatio-temporal Context. The study of context surrounding news events refers to
identify the communities interested in certain events, as well as the events
that impact a certain community. We believe that the study of context is useful
to identify communities being affected by specific types of news, and then being
able to study potential sources of bias, or identify relations between locations
(e.g. international relations) based on these insights. 

User Reaction. The idea of filtering events based on the impact they may cause
on the community can deal with the information overload, by being able to
identify important events based on the activity the users had after reacting to
these events. Emergency situations could be handled more effectively, improve
journalistic coverage, or advertising. By identifying the characteristics of
events that cause certain levels of impact it would be possible to automatically
classify or predict the activity new events will induce, based on other
features, such as content-based or network features. 

Multimedia Content. The study of event content and their characteristics is
meant to provide models for automati- cally summarizing (potentially multimedia)
content and identify the relevant aspects of an event, while considering
specific criteria such as coverage, non-redundancy or diversity. The idea is to
automatically model events to produce comprehensive summaries for humans.
High-level quantitative analysis of events refers to the study of macro
relations of events and their general context. Our hypotheses are that (1)
spatio-temporal models can serve as a base for methodologies to study the
relationships between locations, based on the events that the users from these
locations put their attention; (2) reaction models can help us to discriminate
between events which we would know if they will cause a high impact on society,
from those which do not, and hence understand the characteristics of these
events and how they impact the community; (3) content models are useful for
humans to understand events quickly, and comprehend different points of view
that would be very difficult to perceive if they had to be searched through
potentially many messages. We argue that this kind of analysis is novel in an
automated setting, and the proposed models are useful to achieve this goal. The
literature often focuses on specific types of events, such as emergencies
(floods, earthquakes, sports events, etc.), localized events (a small community
or niches), or memes (atomized pieces of information), and not in global and
general events, but to detect, track and store them. We defend the hypothesis
that the proposed models are useful to discover knowledge about world events in
this setting.


Research Questions 

How to model heterogeneous user-contributed data into representative events? 

To what extent social media information can be rearranged in order to extract
useful knowledge about real-world events? 

How transferable are the proposed methodologies and results from one social
networking service to another? 

What is the lowest level of granularity in which spatio-temporal context can be
measured to identify similar events or communities?

How to define a sound measure of reaction, impact or user activity? 

How to validate specific criteria, such as diversity or coverage, when
evaluating automatic summaries? 

Research Challenges

The identified research challenges are the following: 

Obtain useful and relevant knowledge from microblogs about real-world events.
Social media offers a partial view of the world. Also, mainstream topics
obfuscate distinct points of view, which can obstruct retrieval of diverse
content. Because users are frequently posting messages about their own lives,
daily situations, or general topics, trends can be only visible when looking at
large volumes of data. Besides, users tipically post very short texts,
frequently with spelling and grammar errors. This makes identification of events
and relevant content a very difficult task.


Identify relevant information from potentially noisy data. Due to the
characteristics of Twitter (or any other social networking service), usually
messages are very short and with grammar and spelling errors. Also, users
spontaneously create new ways to refer to the same entities (e.g. via the use of
hashtags, emojis, or abbreviations), which makes it difficult to identify more
relevant content when detecting events.

Take into consideration information of different modalities. Relevant multimedia
content is also difficult to identify. Multimedia content is represented by
images, videos, text, a mixture of them, etc. This information can be exploited
to improve the effectiveness of the proposed methodologies. The challenge comes
in how to identify such content in an efficient way, how to deal with duplicated
or quasi-duplicated content, and how to evaluate the effectiveness of
methodologies when presenting multimedia content. 

Deal with the potential bias in the collected data. As we stated above, social
media offers a partial view of the world. Furthermore, the employed
methodologies to retrieve or identify events from social media may be biased
depending on several factors. For instance, our dataset is collected using news
outlets as sources, being the majority of the outlets coming from the USA or the
UK. This is a huge challenge in order to provide generalizable results from the
proposed methodologies. Also, it is challenging to ensure that our results are
as diverse as the utilized data source. 

General Methodology The general methodology is meant to be applicable to each of
the steps of this work, by being adapted to them. The proposed methodology is as
follows: 

1. Data Collection and Event Idenfitication. 


The event identification step uses a relevance feedback-like procedure to aim
for a high precision and recall when retrieving event-related posts. The data
collection uses several news outlets as sources, as in previous work. These news
outlets publish content about recent news. The assumption is that if a news
event is gaining importance, then several news outlets will report it. Using
standard Data Mining techniques, such as Frequent Itemset Mining, the next step
is to retrieve the most common words shared by these multiple reports from
different sources, marking these words as keywords. These keywords are used as
search terms for querying user-contributed posts about them. These posts
constitute relevant content for the news event. 

2. Feature Extraction. The next step is to extract relevant features from
events. As stated above, we consider an event as a high-level representation of
a real-world occurrence, consisting in all the occurrence-related posts from
different users on the social networking site. With this in mind, the feature
extraction phase obtain characteristics from the collective of information about
events, not necessarily individual posts. For example, when studying context
features, we would like to identify what place a particular event is referring
to, as opposed to identify from where a specific user is publishing content,
which is also a desired feature. 

3. Modeling. The next step is to propose flexible models for events using the
features computed in the previous step. The purpose of the proposed models is to
represent events in a succint way for different applications. This involves a
representation for each one of the perspectives to study: context, reaction and
content. 

4. Exploration, Analysis and Application. From the computed features and the
proposed models, the next part is to explore and analyze the events to get
insights and generate new knowledge. New applications may arise from these
models, for example, prediction models or automatic summarization algorithms. 

5. Validation and Evaluation. The final step in the pipeline is to validate
results, by proposing or using standard 4evaluation metrics, such as precision,
recall, ROUGE [47], comparation with gold standards, or human evaluation. This
involves validation of previous methodologies and evaluation of the proposed
models and applications.


\end{intro}